### 2021-05-01 newvariant.tex
done?

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant conditional on observed relative speed under five epidemiological conditions.}
The relative strength of the new variant $\rho$ conditional on the speed of the wild type $\rw$, the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\rw$ (and therefore $\rv$) are written on top right corners of each panel.
}
\label{fig:relstrength}
\end{figure}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{relspeed.pdf}
\caption{
\textbf{Relative speed of the new variant given assumed relative strength under five epidemiological conditions.}
The relative speed of the new variant $\delta$ given strength of the wild type $\Rw$, ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and squared coefficient of variation in generation intervals $\kappa$.
Relative speed of the new variant $\delta$ is calculated using $\rho=1.61$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\Rw$ (and therefore $\Rv$) are written on top right corners of each panel.
}
\label{fig:relspeed}
\end{figure}

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Methods):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
%% JD: How does this convoluted LaTeX not make you crazy?
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions \citep{flaxman2020Rt}, we model $\Rw(t)$ using a step function (\fref{Rtbias}) and let $\Rv(t) = \rho \Rw(t)$.

\begin{figure}[!pht]
\includegraphics[width=\textwidth]{Rtbias.pdf}
\caption{
\textbf{Estimates of relative strength over time under different scenarios.}
(A, D, G) True (solid lines) and estimated (dashed lines) reproduction numbers of the new variant and the wild type over time.
(B, E, H) True (purple, solid) and estimated (orange, dashed) ratios between reproduction numbers of the new variant and the wild type over time.
(C, F, I) Phase planes (time is implicit) showing true (purple, solid) and estimated (orange, dashed) relationships between estimated reproduction numbers.
Blue dotted lines represent the regression lines of the estimated variant reproduction numbers against the estimated wild type reproduction numbers.
Gray lines represent the one-to-one line.
Top row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type and the variant.
Middle row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type but is longer than that of the variant (4 days).
Bottom row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type but is longer than that of the variant (6 days).
For all simulations, squared coefficient of variation in generation intervals is assumed to equal $\kappa = 1/5$.
}
\label{fig:Rtbias}
\end{figure}

@@

We use a two-strain renewal equation to simulate the spread of the variant and the wild type.
Ignoring birth and death, the incidence of infection caused by the variant $\iv$ and the wild type $\iw$ assuming perfect cross immunity is given by:
\begin{align}
\frac{dS}{dt} &= - \iv(t) - \iw(t)\\
\iv(t) &= \Rv(t) \int_{0}^\infty \iv(t-\tau) g_{\mathrm{var}}(\tau) \dtau\\
\iw(t) &= \Rw(t) \int_{0}^\infty \iw(t-\tau) g_{\mathrm{wt}}(\tau) \dtau
\end{align}
where $S$ represents the proportion of susceptible individuals.
We discretize the model at the time scale of 0.025 days as described in \citep{park2021forward} and simulated the model with $\iv(0) = 0.001$ and $\iw(0) = 0.1$.
We assume that $\Rw(t)$ remains at 2 for 30 days, immediately decreases and stays at 0.5 until day 60, and immediately increases back to 1.2.
Finally, we assume $\Rv(t) = \rho \Rw(t)$.
From simulated incidence curves between day 15 and 70, we estimate the instantaneous reproduction number using \eref{rt}.
We ignore incidence before day 15 to remove any potential transient effects.

>>>>

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant assuming a fixed speed advantage $\delta$ under five epidemiological conditions.}
The relative strength of the new variant $\rho$ conditional on the speed of the wild type $\rw$, the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\rw$ (and therefore $\rv$) are written on top right corners of each panel.
}
\label{fig:relstrength}
\end{figure}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{relspeed.pdf}
\caption{
\textbf{Relative speed of the new variant assuming a fixed strength advantage $\rho$ u under five epidemiological conditions.}
The relative speed of the new variant $\delta$ given strength of the wild type $\Rw$, ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and squared coefficient of variation in generation intervals $\kappa$.
Relative speed of the new variant $\delta$ is calculated using $\rho=1.61$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\Rw$ (and therefore $\Rv$) are written on top right corners of each panel.
}
\label{fig:relspeed}
\end{figure}

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Methods):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions, we let $\Rw(t)$ to decrease from 2 to 0.4 around day 30 and increase back up to 1 around day 60 and assume $\Rv(t) = \rho \Rw(t)$.
Although previous studies have modeled the impact of non-pharmaceutical interventions as a step function \citep{flaxman2020Rt}, we choose to use a smooth function to model $\Rw(t)$ (\fref{Rtbias}; see Methods) given the possibility that behavioral changes may affect transmission before and after interventions take place.
In Supplementary Figure S1, we show that using a step function does not affect our conclusions.

\begin{figure}[!pht]
\begin{center}
\includegraphics[width=0.9\textwidth]{Rtbias_smooth.pdf}
\caption{
\textbf{Estimates of relative strength over time under different scenarios.}
(A, D, G) True (solid lines) and estimated (dashed lines) reproduction numbers of the new variant and the wild type over time.
(B, E, H) True (purple, solid) and estimated (orange, dashed) ratios between reproduction numbers of the new variant and the wild type over time.
(C, F, I) Phase planes (time is implicit) showing true (purple, solid) and estimated (orange, dashed) relationships between estimated reproduction numbers.
Blue dotted lines represent the regression lines of the estimated variant reproduction numbers against the estimated wild type reproduction numbers.
Gray lines represent the one-to-one line.
Top row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type and the variant.
Middle row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type but is longer than that of the variant (4 days).
Bottom row: the assumed mean generation-interval (5 days) is equal to the mean generation-interval of the wild type but is longer than that of the variant (6 days).
For all simulations, squared coefficient of variation in generation intervals is assumed to equal $\kappa = 1/5$.
}
\end{center}
\label{fig:Rtbias}
\end{figure}

@@

We use a two-strain renewal equation to simulate the spread of the variant and the wild type.
Ignoring birth and death, the incidence of infection caused by the variant $\iv$ and the wild type $\iw$ assuming perfect cross immunity is given by:
\begin{align}
\frac{dS}{dt} &= - \iv(t) - \iw(t)\\
\iv(t) &= \Rv(t) \int_{0}^\infty \iv(t-\tau) g_{\mathrm{var}}(\tau) \dtau\\
\iw(t) &= \Rw(t) \int_{0}^\infty \iw(t-\tau) g_{\mathrm{wt}}(\tau) \dtau
\end{align}
where $S$ represents the proportion of susceptible individuals.
We discretize the model at the time scale of 0.025 days as described in \citep{park2021forward} and simulated the model with $\iv(0) = 0.001$ and $\iw(0) = 0.1$.
To allow for smooth changes in $\Rw(t)$ around day 30 and 60, we let:
\begin{equation}
\Rw(t) = 2 \times \left[1 - 0.8 \left(\frac{1}{2} + \frac{\arctan(t-30)}{\pi} \right) \right] \times \left[1 + 1.5 \left(\frac{1}{2} + \frac{\arctan(t-60)}{\pi} \right) \right].
\end{equation}
In this case, $\Rw(t)$ remains around 2 and slowly decreases to $0.4 = 2 \times (1-0.8)$ around day 30, and slowly increases back up to $1 = 0.4 \times (1 + 1.5)$ around day 60.
Finally, we assume $\Rv(t) = \rho \Rw(t)$.
From simulated incidence curves between day 15 and 70, we estimate the instantaneous reproduction number using \eref{rt}.
We ignore incidence before day 15 to remove any potential transient effects.

\pagebreak

\section*{Supplementary Figure}

\setcounter{figure}{0}    
\renewcommand\thefigure{S\arabic{figure}}

\begin{figure}[!pht]
\begin{center}
\includegraphics[width=0.9\textwidth]{Rtbias.pdf}
\caption{
\textbf{Estimates of relative strength over time under different scenarios assuming step changes in reproduction number.}
See Figure 3 of the original text for figure caption.
}
\end{center}
\label{fig:Rtbias}
\end{figure}

\pagebreak
